---
title: "Discussion01 - Calcualting summary statistics"
output:
  html_document:
    df_print: paged
---

# Define Vector

In R a "vector" represents a collection of observation and
is defined by "c(a,b,c,...)" where a,b,c.. represent individual
observations which could be categorical or numeric. We will
stick to numeric observations for this discussion.


```{r}
y <- c(10, 13, 12, 15, 23, 20)
```

The `<-` is equivalent to an "=" sign and indicates that 
we are setting the "value" of `y` to be the observation vector.
You can also use the `=` sign  in place of `<-`, but it is 
usually a good practice to stick to `<-`.


# Median

To calculate median, we simply use the `median` command:

```{r}
median(y)
```

# Mean

Similarly, mean is calculate by the `mean` command:

```{r}
mean(y)
```

# Quantiles

To calculate Q1/Q3, we will use the `quantile` function 

```{r}
quantile(y)
```

# IQR

```{r}
IQR(y)
```

# Standard Deviation

```{r}
sd(y)
```

# Histogram

```{r}
hist(y)
```

# Boxplot

```{r}
boxplot(y)
```