---
title: "Discussion #4"
date: 09/23/2019
---


# Simulating a binomial random variable

R has a lot of utility functions for generating random variables.
We will start by generating a binomial random variable. A binomial
random variable can be generated by `rbinom(n, size, probability_of_sucess)` command.
Note that `n` here refers to how many such random variables we want to generate.
You can generate 1, but that is never sufficient to study anything, so we will
generate 1000 of such random variables.

## N = 10, p = 0.1

We will first simulate a binomial with probability of success 0.1 and when number of trials (size) is 10. We will simulate 1000 such random numbers.

```{r}
x <- rbinom(1000, 100, 0.1)
```

Let's plot it to visualize it:

```{r}
hist(x, main="N=Binomial 10,p=0.1")
```
## Question: Where is the mean located? Is it easy to identify?


## N = 100, p = 0.1

Let's increase the number of trials to 50, keeping the probability constant.
```{r}
x <- rbinom(1000, 100, 0.1)
hist(x, main="Binomial N=100, p=0.1")
```

## Question: Can you locate the mean now with N=100, p=0.1?


## N=100, p=0.5

Let's increase the success probability to 0.5
```{r}
x <- rbinom(1000, 100, 0.5)
hist(x, main="Binomial N=100, p=0.5")
```

## Question: Is there some other type of distribution you can think of that is approximately similar to the binomial random variable?



## Creating a normal random variable.

The mean of a biniomial random variable with $n$ trials and success of probability $p$ is $np$ and its variance is given by $np(1-p)$.

For $n=100, p=0.5$, the mean and variance are $50$ and $25$ respectively.
Let's try to create a Normal random variable with the $\mu=50$ and $\sigma^2=25$. 
This can be done again by using a similar command that we used for generating
random variable: `rnorm(n, mean, var)`. Recall that you can always do "?rnorm"
or "help(rnorm)" to lookup its usage and examples.


## Normal (50,25)

```{r}
x <- rnorm(1000, 50, 25)
hist(x, main="Normal(50,25)")
```


## Checking for normality of a binomial random variable

A QQ plot can be used to check if a binomial random variable is approximately normal.
Plotting a QQ plot is very easy in R using `qqnorm(X)`.

We first test it on a Normal distribution:

```{r}
x.normal <- rnorm(1000, 50, 25)
qqnorm(x.normal,  main="QQ plot of  Normal(50,25)")
qqline(x.normal)
```

And we finally repeat it for the binomial distribution:

```{r}
y.binomial <- rbinom(1000, 100, 0.5)
qqnorm(y.binomial,  main="QQ plot of  Binomial N=100,p=0.5")
qqline(y.binomial)
```