---
title: "Discussion 07"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "10/14/2019"
---

# VitaminC vs Placebo Example

## Calculate everything as if by hand

```{r}
vitaminC <- c(5.5, 6.0, 7.0, 6.0, 7.5, 6.0, 7.5, 5.5, 7.0, 6.5)
placebo <- c(6.5, 6.0, 8.5, 7.0, 6.5, 8.0, 7.5, 6.5, 7.5, 6.0, 8.5, 7.0)

mean.vitaminC <- mean(vitaminC)
mean.placebo <- mean(placebo)

n.vitaminC <- length(vitaminC)
n.placebo <- length(placebo)

sd.vitaminC <- sd(vitaminC)
sd.placebo <- sd(placebo)

se.vitaminC <- sd.vitaminC/sqrt(n.vitaminC)
se.placebo <- sd.placebo/sqrt(n.placebo)

se.difference <- sqrt(se.vitaminC^2 + se.placebo^2)
tstat <- (mean.vitaminC-mean.placebo)/(se.difference)

```

For calculating the degrees of freedom we will use the 'exact' formula
as in Equation 6.7.1 in the textbook:

$$
df = \frac{(SE_1^2+SE_2^2)^2}{\frac{SE_1^4}{n_1-1} + \frac{SE_2^4}{n_2-1}}
$$

```{r}

df <- (se.vitaminC^2+se.placebo^2)^2/((se.vitaminC^4)/(n.vitaminC-1)+(se.placebo^4)/(n.placebo-1))
```

Now we have tstat, say '$t$' and the degrees of freedom '$df$', the two-sided pvalue can be calculated using:

$$
pvalue = 2 \times P(T \leq -|t|) = 2 \times P(T \geq |t|)
$$
Here, $T$ represents the t-distribution, and $|x|$ indicates the absolute value of $x$. You can approximately calculate the range this p-value will lie using Table 4 in the textbook as follows:

1. Lookup the t-value corresponding to degrees of freedom $20$ such that $|t|=|-1.9253| = 1.9253. The two such entries are 1.850 and 2.00 corresponding to an upper tail probability values for $0.04$ and $0.03$. 

2. Since Table 4 only lists the upper tail probability while we are interested
in two-sided p-value, we can make it two sided by multiplying by $2$. This is possible since $t$ distribution is symmetric and as such the area on the lower tail is equal to the area in the upper tail.

The approixmate p-value using this would be $0.03 \times 2 = 0.06$.

## Using R:

Given that we have the t-statistic `tstat` calculated above and `df`, we can also use the `pt` function from R to caclualte the p-value more accurately: \texttt{p-value = 2*pt(-abs(t), df)}

```{r}
tstat
```

```{r}
df
```

```{r}
pvalue.calculated <- 2 * pt(-abs(tstat), df)
pvalue.calculated
```

0.06856691 is expected from our approximation study as well.


## Calculating 95% confidence interval

Given that we have the degrees of freedom `df`, we can calculate the 95% confidence interval given by:

$$
(\mu_1-\mu_2) \pm t_{0.025}SE_{\bar{Y}_1-\bar{Y}_2}
$$
To calculate $t_{0.025}$ we will make use of the `qt` function in R which gives us quantiles:

```{r}
t <- qt(0.05/2, df)
upper.ci <- (mean.vitaminC-mean.placebo) + se.difference * t
lower.ci <- (mean.vitaminC-mean.placebo) - se.difference * t
```

```{r}
upper.ci
```

```{r}
lower.ci
```

## t-test using R

We can also simplify everything using R:

```{r}
t.test(vitaminC, placebo)
```


Look through the output and check if all the values we calculated matched
