---
title: "Discussion 06"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

# Central Limit Theorem

Accoruding to the central limit theorem, the sample mean is normally
distributed. Note that this holds irrespective of the distribution
defining the population. The population itself need not follow a
normal distribution.

## Simulate a population that is not distributed randomly
```{r}
population1 <- rnorm(20000, mean = 10, sd = 3)
population2 <- rnorm(80000, mean = 70, sd = 10)

# Create a population of 100,000 numbers
population <- c(population1, population2)
hist(population)
```

Clearly the above population is not normally distributed.
We will now sample this population for different number of times
and see its distribution.

## Simulating Sample Means

We will start of with sampling this population and calculate the sample mean. The
sample size is set to two and we obtain 100 such samples. Does the
distribution look normal to you?

```{r}
sample.size <- 2
number.samples <- 100
# Obtain 100 samples each of size 2 by sampling the population
ybars <- replicate(number.samples, mean(sample(population, sample.size,replace = TRUE)))
hist(ybars)
```


## Let's increase the sample size to 10:

```{r}
sample.size <- 10
number.samples <- 100
ybars <- replicate(number.samples, mean(sample(population, sample.size, replace = TRUE)))
hist(ybars)
```

## What happens if we increase the sample size to 30?

```{r}
sample.size <- 30
number.samples <- 100
ybars <- replicate(number.samples, mean(sample(population, sample.size, replace = TRUE)))
hist(ybars)
```





## What if we sample the population for samples of size 30 and look at its distribution (not its mean), would it be normally distributed?

```{r}
sample.size <- 30
number.samples <- 100
y <- replicate(number.samples, sample(population, sample.size,replace = TRUE))
hist(y)
```

# Aside 1: Calculating t statistic in R

You can use R to calculate t-statistic in using the `qt` (qunatile of t) command.
For example, to calculate the t-statistic cooresponding to 95% confidence interval
with 12.8 degrees of freedom:

```{r}
abs(qt(0.05/2, 12.8))
```

# Aside 2: Calculating z statistic in R

Similarly, z-statistic can be obtained by using the `qnorm` (quantile of normal) command.
Z score corresponding to $P(Z<0.05)$ is given as:

```{r}
qnorm(0.05)
```


